Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-21 ms/SI.Rmd
Use 'nothotaxa' instead of 'hybrid'

**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

**`r s_figure("apo-genus")`**. Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus.

>>>>

**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

**`r s_figure("apo-genus")`**. Rate of apomixis in genera Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus. Excludes non-native taxa and nothotaxa.

### 2021-09-21 ms/manuscript.Rmd
Use 'nothotaxa' instead of 'hybrid'

**Methods:** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. 
We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. 
We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
We compared the overlap of significantly diverse grid-cells with existing conservation zones and range maps of native Japanese deer to assess conservation status and threats.

@@

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b]. 
One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].

@@

We used a list of all native, non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
Points (latitude and longitude of each occurrence record) were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].

@@

The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families. 
The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016]. 
The age of the fern clade is estimated to be `r phylogeny_age` ma.

@@

One element of biodiversity that we did not consider here is hybrids. 
The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa. 
Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989]. 
Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
Methods for calculating PD would need to be modified to take into account multiple lineages contributing to one terminal taxon.

@@

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

>>>>

**Methods:** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` taxa. 
We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. 
We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
We compared the overlap of significantly diverse grid-cells with existing conservation zones and range maps of native Japanese deer to assess conservation status and threats.

@@

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b]. 
One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa (excluding nothotaxa) [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].

@@

We used a list of fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
Points (latitude and longitude of each occurrence record) were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].

@@

The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native ferns excluding nothotaxa [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families. 
The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016]. 
The age of the fern clade is estimated to be `r phylogeny_age` ma.

@@

One element of biodiversity that we did not consider here is nothotaxa (hybrids). 
The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa. 
Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989]. 
Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
Methods for calculating PD would need to be modified to take into account multiple lineages contributing to one terminal taxon.

@@

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
